---
title: 'STA 360/602 Homework 1: Data Wrangling in \texttt{R}'
author: "Chavez Cheong"
date: "1/5/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Working with data

1.  (22 points total, equally weighted) The data set **rnf6080.dat** records hourly rainfall at a certain location in Canada, every day from 1960 to 1980.

a. Load the data set into R and make it a data frame called \texttt{rain.df}.  What command did you use?

```{r load-data}
rain.df <- read.table('data/rnf6080.dat')
```

I used the `read.table()` command.

b. How many rows and columns does \texttt{rain.df} have?  How do you know?  (If there are not 5070 rows and 27 columns, you did something wrong in the first part of the problem.)

```{r dataset-dimension}
dim(rain.df)
```
\texttt{rain.df} has 5070 rows and 27 columns. I know this because I can see the dimensions of the data frame with the `dim()` command.

c. What command would you use to get the names of the columns of \texttt{rain.df}? What are those names?

```{r}
names(rain.df)
```

I would use the `names` command to get the names of the coloumns of \texttt{rain.df}. These names are V1 through V27 (V followed by the corresponding column number).

d. What command would you use to get the value at row 2, column 4? What is the value?

```{r}
rain.df[2,4]
```
I would use the `[]` command, specifying the row number as the first element and the column number as the second element. The value is 0.

e. What command would you use to display the whole second row?  What is the content of that row?

```{r}
rain.df[2,]
```

I would use the `[]` command again, this time specifying 2, the first element as the row number, but leaving the column number blank as the second element, so that I display the whole second row. The content of that row is as displayed above.

f.  What does the following command do?
```{r}
names(rain.df) <- c("year","month","day",seq(0,23))
```

This command renames the first three columns as "year", "month" and "day", and then renames columns "V4" to "V27" as "0" through to "23".

g.  Create a new column called \texttt{daily}, which is the sum of the 24 hourly columns.

```{r}
rain.df$daily <- rowSums(rain.df[(4:27)])
```

h. Give the command you would use to create a histogram of the daily rainfall amounts.  Please make sure to attach your figures in your .pdf report.

```{r}
hist(rain.df$daily, main = 'Histogram of Daily Rainfall', xlab = "Daily Rainfall")
```
i. Explain why that histogram above cannot possibly be right.

It is not possible that there are days with negative rainfall.

j. Give the command you would use to fix the data frame.

```{r}
rain.df[rain.df < 0] <- 0
```

I would use the `[rain.df < 0]` command to replace all negative values in the dataframe with 0. 

k. Create a corrected histogram and again include it as part of your submitted report. Explain why it is
more reasonable than the previous histogram.

```{r}
hist(rain.df$daily, main = 'Histogram of Daily Rainfall', xlab = "Daily Rainfall")
```

Now there are no days with negative total daily rainfall, and all values are positive values, which is more reasonable than the previous histogram which included negative values.

***Data types***

2. (9 points, equally weighted) Make sure your answers to different parts of this problem are compatible with each other.

a. For each of the following commands, either explain why they should be errors, or explain the non-erroneous result.
```
x <- c("5","12","7")
max(x)
sort(x)
sum(x)
```

`max(x)` would return "7", as between all the three first characters of each element ("1" for "12", "5" for "5" and "7" for "7"), "7" is the largest lexicographical value.

`sort(x)` would return the vector ("12", "5", "7"), as it sorts the elements by the lexicographic order of the first character of each element ("1" for "12", "5" for "5" and "7" for "7").

`sum(x)` would throw an error, as all elements in the vector `x` are characters, and the `sum` function only takes numeric arguments.

b. For the next two commands, either explain their results, or why they
should produce errors.
```
y <- c("5",7,12)
y[2] + y[3]
```

The command `y <- c("5",7,12)` assigns the vector ("5", "7", "12") to the variable y. This is because all elements in a vector must be of the same type, so R automatically converted the integers 7 and 12 into characters. 

`y[2] + y[3]` will return an error, as the `+` binary operator only accepts numeric arguments, but y[2] and y[3] store "7" and "12", respectively, which are characters and non-numeric.

c. For the next two commands, either explain their results, or why they should
produce errors.
```
z <- data.frame(z1="5",z2=7,z3=12)
z[1,2] + z[1,3]
```

`z <- data.frame(z1="5",z2=7,z3=12)` will assign a


3. (3 pts, equally weighted). 

a.) What is the point of reproducible code? 



b.) Given an example of why making your code reproducible is important for you to know in this class and moving forward. 



c.) On a scale of 1 (easy) -- 10 (hard), how hard was this assignment. If this assignment was hard ($>5$), please state in one sentence what you struggled with. 

This assignment is a 2 in terms of difficulty.